Supporting Information: Time Series Modeling of Solute Transport in an H\textsubscript{II} Phase Lyotropic Liquid Crystal Membrane

### 2019-12-12 Ben_Manuscripts/stochastic_transport/Draft.tex
Figure updates

  We determined the rate at which solutes cross macroscopic-length pores based on the
  Hill relation:~\cite{hill_free_1989}
  \begin{equation}
  J = \frac{1}{MFPT}
  \label{eqn:hill_relation}
  \end{equation}
  where $J$ is the single particle solute flux and MFPT refers to the mean first passage
  time. 
  %MRS5: I would phrase below: 
  To account for input concentration dependence of the flux, assuming that 
  particles are uncorrelated, 
  %For particle concentrations higher than 1 at the pore entrance, 
  one can multiply
  Equation~\ref{eqn:hill_relation} by the total number of particles to get the total flux.
  In the context of our work, the MFPT describes the average length of time it takes a particle
  to move from the pore entrance to the pore exit.

@@

  To come up with an analytical equation describing the passage time distributions, 
  one can frame the problem as a pulse of particles instantaneously released at the 
  pore inlet ($z=0$) which moves at a constant velocity and spreads out as it 
  approaches $L$. This gives results equivalent to if we had released each particle
  individually and then analyzed the positions of the ensemble of trajectories as 
  a function of time. 
  %MRS5: I would put this next section up until just before Eq. 15 in the supporting information (keep any citations in the main text). 
  To derive an analytical equation, first consider an initial pulse spreading out 
  over time with a fixed mean. We can solve for the time-dependent probability density
  of particle positions by solving the one dimensional diffusion equation:
  \begin{equation}
  \frac{\partial p}{\partial t} = D \frac{\partial^2 p}{\partial z^2}
  \end{equation}
  The appropriate initial and boundary conditions are:
  $$BC1: t > 0, z = \infty, p = 0$$
  $$BC2: t > 0, z = 0, \frac{\partial p}{\partial z} = 0$$
  $$IC: t = 0, c = M\delta(z)$$
  where $M$ is the total number of particles in the system. It has been shown 
  elsewhere that the solution to this equation is:~\cite{cussler_diffusion:_2009}
  \begin{equation}
  p(z, t) = \frac{1}{\sqrt{4 \pi D t}}\exp\bigg(\frac{-z^2}{4Dt}\bigg)
  \end{equation}
  We can make the substitution $z = z - vt$, where $v$ represents a constant average
  velocity, in order to linearly shift the mean as a function of time:
  \begin{equation}
  p(z, t) = \frac{1}{\sqrt{4 \pi D t}}\exp\bigg(\frac{-(z - vt)^2}{4Dt}\bigg)
  \end{equation}
  One can track the fraction of particles, $F$, that have crossed the pore 
  boundary by integrating:
  \begin{equation}
  F(t) = \int_L^\infty p~dz = \mathrm{erfc}\bigg(\frac{L - vt}{2\sqrt{D t}}\bigg)
  \end{equation}
  where $L$ is the pore length. This represents the cumulative first passage 
  time distribution so we take its derivative in order to arrive at the first
  passage time distribution:
  \begin{equation}
  P(t) = -\frac{1}{\sqrt{\pi}}e^{-(L - vt)^2 / (4Dt)}\bigg(-\frac{D(L - vt)}{4(Dt)^{3/2}} - \frac{v}{2\sqrt{Dt}}\bigg)
  \label{eqn:passage_times}
  \end{equation} 
  where the only free parameters for fitting are $v$ and $D$. We calculated the
  expected value of Equation~\ref{eqn:passage_times} in order to get the MFPT.

  We used the ratio of solute fluxes in order to determine membrane selectivity, $S_{ij}$,
  towards solutes. Selectivity of solute $i$ versus $j$ is defined in terms of the
  ratio of solute permeabilities, $P$:~\cite{guo_pervaporation_2004}
  \begin{equation}
  S_{ij} = \frac{P_i}{P_j}
  \end{equation}
  We can relate this to solute flux using Kedem and Katchalsky's equations for 
  solvent volumetric flux, $J_v$, and solute flux, $J_s$:~\cite{kedem_permeability_1963}
  %al-zoubi_rejection_2007 clarifies that these equations can be used for nanofiltration, ultrafiltration etc. by treating the membrane as a black box subject to irreversible thermodynamics.
  \begin{equation}
  J_v = L_p(\Delta P - \sigma\Delta \pi)
  \end{equation} 
  \begin{equation}
  J_s = P_s \Delta C + (1 - \sigma)CJ_v
  \label{eqn:solute_flux}
  \end{equation}
  where $L_p$ is the pure water permeability, $\Delta P$ and $\Delta \pi$ are the 
  trans-membrane hydraulic and osmotic pressure differences, $\sigma$ is the reflection
  coefficient, $P_s$ is the solute permeability, $\Delta C$ is the trans-membrane
  solute concentration difference and $C$ is the mean solute concentration. Since our
  simulations do not include convective solute flux, we eliminate the second term 
  of Equation~\ref{eqn:solute_flux} which allows us to derive a simple expression
  for selectivity in terms of solute flux:
  \begin{equation}
  S_{ij} = \frac{J_i / \Delta C_i}{J_j / \Delta C_j}
  \label{eqn:selectivity}
  \end{equation}

@@

  %BJC4: not sure if the next couple paragraphs belong in methods 
  %MRS5: this could potentially be in methods.  Alternatives can be discussed there.
  It is computationally infeasible to simulate trajectories long enough that they
  traverse the length of a macroscopic pore. To date, the thinnest H\textsubscript{II}
  LLC membrane synthesized with the monomer in this work was 7$\mu$m thick. Using
  24 cores to simulate trajectory realizations in parallel, it takes on the order 
  of 1 day to simulate 50000 realizations of solutes traversing a 50 nm pore. The 
  RAM requirements and performance scales greater than linearly and thus would 
  take an infeasible amount of memory and time to simulate transport through a 
  pore over 100 times longer.

  %BJC4: methods?  
  %MRS5: This requires knowing the data, so maybe here. 
  We used simulated trajectories which traverse computationally-reasonable length
  pores in order to construct an empirical model which one can use to estimate 
  particle flux for arbitrary length pores. As shown in Figure~\ref{fig:flux_curves},
  the flux appears to scale according to a power law of the form:
  \begin{equation}
  J(L) = cL^{-\beta} 
  \label{eqn:flux_decay}
  \end{equation}
  %while the MFPT scales as its inverse, $(1/c)L^{\beta}$.

@@

  %BJC: need simulations with increased length for L=35-50 nm pores of acetic acid and methanol
  %BJC: see of clauset has better way of fitting t^-alpha (alpha might be too big)
  %BJC: not sure mfpt is necessary. Takes up space.
  %BJC: change alpha to beta
  %BJC: make a chart with parameters
  %BJC: should have a paragraph about c -- roughly follows their MSDs.
  \begin{figure}
  \centering
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{flux_curves.pdf}
  \caption{Fits to $cL^{-\beta}$}\label{fig:flux_curves_regular}
  \end{subfigure}
  % BJC: could use better color scheme. Maybe outline hurst parameters
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{flux_parameters.pdf}
  \caption{}\label{fig:flux_parameters}
  \end{subfigure}
%  \begin{subfigure}{0.475\textwidth}
%  \includegraphics[width=\textwidth]{mfpt_curves.pdf}
%  \caption{}\label{fig:mfpt_curves}
%  \end{subfigure}
  %BJC: going out to 3 decimal places seems silly
  %BJC: might get closer to beta=2 if I use different curve fitting technique than fitting line to log-log
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{flux_curves_brownian.pdf}
  \caption{Fixed $H=0.5$}\label{fig:flux_curves_brownian}
  \end{subfigure}
%  \begin{subfigure}{0.475\textwidth}
%  \includegraphics[width=\textwidth]{mfpt_curves_brownian.pdf}
%  \caption{}\label{fig:mfpt_curves_brownian}
%  \end{subfigure}
  %MRS5: the b) pane is a little messy.  If c) is moved to the supporting, maybe can make b into two figures, and 
  % have each just show 1 point? 
  \caption{(a) The single particle solute flux decays with increasing pore length. 
  Solutes with high MSDs have higher fluxes. (b) We fit the single particle solute flux 
  versus pore length, $L$, to a power law function of the form $cL^{-\beta}$ (dashed 
  lines in (a)). Solutes with higher fluxes have higher values of $c$. Solutes with 
  smaller Hurst parameters, $H$, have larger values of $\beta$. (c) If we remove 
  correlation between hops (set $H$=0.5), the flux increases significantly and 
  $\beta \sim 2$ for all solutes independent of dwell distribution
  parameters.}\label{fig:flux_curves}
  \end{figure}

@@

  Of course, this approach to modeling LLC membrane selectivity neglects a number of
  factors that influence real-world aqueous membrane separations. The most glaring is 
  that we neglected convective flux which should work to increase overall solute flux 
  and lower selectivity (see Equation ~\ref{eqn:solute_flux}). Defects in the pore 
  structure will have a strong influence on solute flux which may exacerbate the length 
  dependence of flux. Even in a perfectly structured membrane, the barrier to pore
  entry from the bulk solution will have its own influence on experimentally observed 
  selectivities. This energetic barrier can be made even worse by foulants deposited on
  the membrane's surface. Therefore, our predictions should be primarily used to 
  identify the most promising types of selective separations for an H\textsubscript{II}
  LLC membrane made from a specific monomer.
  %BJC: though about mentioning fouli

  %MRS5: sentence fragment left here?
    Neglects concentration polarization, membrane defects and fouling.

>>>>

  We determined the rate at which solutes cross macroscopic-length pores based on the
  Hill relation:~\cite{hill_free_1989}
  \begin{equation}
  J = \frac{1}{MFPT}
  \label{eqn:hill_relation}
  \end{equation}
  where $J$ is the single particle solute flux and MFPT refers to the mean first passage
  time. To account for input concentration dependence of the flux, assuming that particles
  are uncorrelated, one can multiply Equation~\ref{eqn:hill_relation} by the total number
  of particles to get the total flux. In the context of our work, the MFPT describes the
  average length of time it takes a particle to move from the pore entrance to the pore exit.

@@

  To come up with an analytical equation describing the passage time distributions, 
  one can frame the problem as a pulse of particles instantaneously released at the 
  pore inlet ($z=0$) which moves at a constant velocity, $v$, and spreads out as it 
  approaches $L$. This spreading is parameterized by an effective diffusivity 
  parameter, $D$. This approach gives results equivalent to if we had released each
  particle individually and then analyzed the positions of the ensemble of trajectories as 
  a function of time. The analytical expression describing the distribution of 
  first passage times is:~\cite{cussler_diffusion:_2009}
  \begin{equation}
  P(t) = -\frac{1}{\sqrt{\pi}}e^{-(L - vt)^2 / (4Dt)}\bigg(-\frac{D(L - vt)}{4(Dt)^{3/2}} - \frac{v}{2\sqrt{Dt}}\bigg)
  \label{eqn:passage_times}
  \end{equation} 
  where the only free parameters for fitting are $v$ and $D$. A derivation
  of Equation~\ref{eqn:passage_times} is given in Section~\ref{S-section:fpt_derivation}
  of the Supporting Information. We calculated the expected value of 
  Equation~\ref{eqn:passage_times} in order to get the MFPT.

  %MRS5: I would put this next section up until just before Eq. 15 in the supporting information (keep any citations in the main text). 
  %BJC5: moved to supporting
%  To derive an analytical equation, first consider an initial pulse spreading out 
%  over time with a fixed mean. We can solve for the time-dependent probability density
%  of particle positions by solving the one dimensional diffusion equation:
%  \begin{equation}
%  \frac{\partial p}{\partial t} = D \frac{\partial^2 p}{\partial z^2}
%  \end{equation}
%  The appropriate initial and boundary conditions are:
%  $$BC1: t > 0, z = \infty, p = 0$$
%  $$BC2: t > 0, z = 0, \frac{\partial p}{\partial z} = 0$$
%  $$IC: t = 0, c = M\delta(z)$$
%  where $M$ is the total number of particles in the system. It has been shown 
%  elsewhere that the solution to this equation is:~\cite{cussler_diffusion:_2009}
%  \begin{equation}
%  p(z, t) = \frac{1}{\sqrt{4 \pi D t}}\exp\bigg(\frac{-z^2}{4Dt}\bigg)
%  \end{equation}
%  We can make the substitution $z = z - vt$, where $v$ represents a constant average
%  velocity, in order to linearly shift the mean as a function of time:
%  \begin{equation}
%  p(z, t) = \frac{1}{\sqrt{4 \pi D t}}\exp\bigg(\frac{-(z - vt)^2}{4Dt}\bigg)
%  \end{equation}
%  One can track the fraction of particles, $F$, that have crossed the pore 
%  boundary by integrating:
%  \begin{equation}
%  F(t) = \int_L^\infty p~dz = \mathrm{erfc}\bigg(\frac{L - vt}{2\sqrt{D t}}\bigg)
%  \end{equation}
%  where $L$ is the pore length. This represents the cumulative first passage 
%  time distribution so we take its derivative in order to arrive at the first
%  passage time distribution:

  We used the ratio of solute fluxes in order to determine membrane selectivity, $S_{ij}$,
  towards solutes. Selectivity of solute $i$ versus $j$ is defined in terms of the
  ratio of solute permeabilities, $P$:~\cite{guo_pervaporation_2004}
  \begin{equation}
  S_{ij} = \frac{P_i}{P_j}
  \end{equation}
  We can relate this to solute flux using Kedem and Katchalsky's equations for 
  solvent volumetric flux, $J_v$, and solute flux, $J_s$:~\cite{kedem_permeability_1963,al-zoubi_rejection_2007}
  %al-zoubi_rejection_2007 clarifies that these equations can be used for nanofiltration, ultrafiltration etc. by treating the membrane as a black box subject to irreversible thermodynamics.
  \begin{equation}
  J_v = L_p(\Delta P - \sigma\Delta \pi)
  \end{equation} 
  \begin{equation}
  J_s = P_s \Delta C + (1 - \sigma)CJ_v
  \label{eqn:solute_flux}
  \end{equation}
  where $L_p$ is the pure water permeability, $\Delta P$ and $\Delta \pi$ are the 
  trans-membrane hydraulic and osmotic pressure differences, $\sigma$ is the reflection
  coefficient, $P_s$ is the solute permeability, $\Delta C$ is the trans-membrane
  solute concentration difference and $C$ is the mean solute concentration. Since our
  simulations do not include convective solute flux, we eliminate the second term 
  of Equation~\ref{eqn:solute_flux} which allows us to derive a simple expression
  for selectivity in terms of solute flux:
  \begin{equation}
  S_{ij} = \frac{J_i / \Delta C_i}{J_j / \Delta C_j}
  \label{eqn:selectivity}
  \end{equation}

@@

  %BJC4: not sure if the next couple paragraphs belong in methods 
  %MRS5: this could potentially be in methods.  Alternatives can be discussed there.
  %BJC5: I tried putting this in methods, but it feels weird without stating the
  % empirical equation. And I'm avoiding stating the empirical model in the methods
  % since it is informed by the data.
  It is computationally infeasible to simulate trajectories long enough that they
  traverse the length of a macroscopic pore. To date, the thinnest H\textsubscript{II}
  LLC membrane synthesized with the monomer in this work was 7$\mu$m thick. Using
  24 cores to simulate trajectory realizations in parallel, it takes on the order 
  of 1 day to simulate 50000 realizations of solutes traversing a 50 nm pore. The 
  RAM requirements and performance scales greater than linearly and thus would 
  take an infeasible amount of memory and time to simulate transport through a 
  pore over 100 times longer.

  We used simulated trajectories which traverse computationally-reasonable length
  pores in order to construct an empirical model which one can use to estimate 
  particle flux for arbitrary length pores. As shown in Figure~\ref{fig:flux_curves},
  the flux appears to scale according to a power law of the form:
  \begin{equation}
  J(L) = cL^{-\beta} 
  \label{eqn:flux_decay}
  \end{equation}
  %while the MFPT scales as its inverse, $(1/c)L^{\beta}$.

@@

  %BJC: need simulations with increased length for L=35-50 nm pores of acetic acid and methanol
  %BJC: see of clauset has better way of fitting t^-alpha (alpha might be too big)
  %BJC: not sure mfpt is necessary. Takes up space.
  %BJC: change alpha to beta
  %BJC: make a chart with parameters
  %BJC: should have a paragraph about c -- roughly follows their MSDs.
  \begin{figure}
  \centering
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{flux_curves.pdf}
  \caption{Fits to $cL^{-\beta}$}\label{fig:flux_curves_regular}
  \end{subfigure}
  % BJC: could use better color scheme. Maybe outline hurst parameters
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{c_parameters.pdf}
  \caption{}\label{fig:c_parameters}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{beta_parameters.pdf}
  \caption{}\label{fig:beta_parameters}
  \end{subfigure}
%  \begin{subfigure}{0.475\textwidth}
%  \includegraphics[width=\textwidth]{mfpt_curves.pdf}
%  \caption{}\label{fig:mfpt_curves}
%  \end{subfigure}
  %BJC: going out to 3 decimal places seems silly
  %BJC: might get closer to beta=2 if I use different curve fitting technique than fitting line to log-log
%  \begin{subfigure}{0.325\textwidth}
%  \includegraphics[width=\textwidth]{flux_curves_brownian.pdf}
%  \caption{Fixed $H=0.5$}\label{fig:flux_curves_brownian}
%  \end{subfigure}
%  \begin{subfigure}{0.475\textwidth}
%  \includegraphics[width=\textwidth]{mfpt_curves_brownian.pdf}
%  \caption{}\label{fig:mfpt_curves_brownian}
%  \end{subfigure}
  %MRS5: the b) pane is a little messy.  If c) is moved to the supporting, maybe can make b into two figures, and 
  % have each just show 1 point? 
  \caption{(a) The single particle solute flux decays with increasing pore length. 
  Solutes with high MSDs have higher fluxes. (b) We fit the single particle solute flux 
  versus pore length, $L$, to a power law function of the form $cL^{-\beta}$ (dashed 
  lines in (a)). Solutes with higher fluxes have higher values of $c$. Solutes with 
  smaller Hurst parameters, $H$, have larger values of $\beta$. (c) If we remove 
  correlation between hops (set $H$=0.5), the flux increases significantly and 
  $\beta \sim 2$ for all solutes independent of dwell distribution
  parameters.}\label{fig:flux_curves}
  \end{figure}

@@

  Of course, this approach to modeling LLC membrane selectivity neglects a number of
  factors that influence real-world aqueous membrane separations. The most glaring is 
  that we neglected convective flux which should work to increase overall solute flux 
  and lower selectivity (see Equation ~\ref{eqn:solute_flux}). Defects in the pore 
  structure will have a strong influence on solute flux which may exacerbate the length 
  dependence of flux. Even in a perfectly structured membrane, the barrier to pore
  entry from the bulk solution will have its own influence on experimentally observed 
  selectivities. This energetic barrier can be made even worse by foulants deposited on
  the membrane's surface. Therefore, our predictions should be primarily used to 
  identify the most promising types of selective separations for an H\textsubscript{II}
  LLC membrane made from a specific monomer.

### 2019-12-12 Ben_Manuscripts/stochastic_transport/Supporting_Information.tex
Figure updates

@@

>>>>

  \section{Derivation of Passage Time Distributions}\label{section:fpt_derivation}

  To derive an analytical equation for the mean first passage time as calculated
  in the main text, first consider an initial pulse spreading out over time with
  a fixed mean. We can solve for the time-dependent probability density of 
  particle positions by solving the one dimensional diffusion equation:
  \begin{equation}
  \frac{\partial p}{\partial t} = D \frac{\partial^2 p}{\partial z^2}
  \end{equation}
  The appropriate initial and boundary conditions are:
  $$BC1: t > 0, z = \infty, p = 0$$
  $$BC2: t > 0, z = 0, \frac{\partial p}{\partial z} = 0$$
  $$IC: t = 0, c = M\delta(z)$$
  where $M$ is the total number of particles in the system. It has been shown 
  elsewhere that the solution to this equation is:~\cite{cussler_diffusion:_2009}
  \begin{equation}
  p(z, t) = \frac{1}{\sqrt{4 \pi D t}}\exp\bigg(\frac{-z^2}{4Dt}\bigg)
  \end{equation}
  We can make the substitution $z = z - vt$, where $v$ represents a constant average
  velocity, in order to linearly shift the mean as a function of time:
  \begin{equation}
  p(z, t) = \frac{1}{\sqrt{4 \pi D t}}\exp\bigg(\frac{-(z - vt)^2}{4Dt}\bigg)
  \end{equation}
  One can track the fraction of particles, $F$, that have crossed the pore 
  boundary by integrating:
  \begin{equation}
  F(t) = \int_L^\infty p~dz = \mathrm{erfc}\bigg(\frac{L - vt}{2\sqrt{D t}}\bigg)
  \end{equation}
  where $L$ is the pore length. This represents the cumulative first passage 
  time distribution so we take its derivative in order to arrive at the first
  passage time distribution:
  \begin{equation}
  P(t) = -\frac{1}{\sqrt{\pi}}e^{-(L - vt)^2 / (4Dt)}\bigg(-\frac{D(L - vt)}{4(Dt)^{3/2}} - \frac{v}{2\sqrt{Dt}}\bigg)
  \label{eqn:passage_times}
  \end{equation} 
  where the only free parameters for fitting are $v$ and $D$. We calculated the
  expected value of Equation~\ref{eqn:passage_times} in order to get the MFPT.

@@

  \begin{figure}
  \includegraphics[width=0.5\textwidth]{flux_curves_brownian.pdf}
  \caption{Fixed $H=0.5$}\label{fig:flux_curves_brownian}
  \end{figure}
